Tulai

### 2023-09-23 paper/paper.qmd

```{r}
# Read the data
tl <- read_csv(here("data/Tulai_Lithic_Assamblage.csv"))

# this file is prepared from the google sheet here
# https://docs.google.com/spreadsheets/d/1Q0QZESk412ZQLE24yPs6Rg-7Y9OMX63DHoXeIPNAIFM/edit#gid=0
# and simplified using code in the scraps.R file
excavation_data <- read_csv(here("data/excavation_data.csv"))
```

@@

# BM: this is table 1 in the manuscript
knitr::kable(summary_table_area_filtered,
             caption = "Summary of lithics in the main areas")

@@

@@

# Create the plot: WABI_Level_logscale
# here I do not consider those levels with a value 0 of retouch frequency. 
filtered_data <- 
  sum_depth_sheet %>% 
  filter(lithic_density > 0 & retouch_frequency > 0)

WABI_Level <- ggplot(filtered_data,  
            aes(x = lithic_density,
                y = retouch_frequency)) +
  geom_point(color = "black", 
             size = 1.5) +
  geom_smooth(method = "lm", 
              se = TRUE, 
              color = "blue", 
              linewidth = 0.5) +
  # Add text labels based on 'label' column
  geom_text_repel(aes(label = label2)) +
  stat_poly_eq(aes(label = paste(stat(eq.label), 
                                  stat(rr.label), 
                                  stat(p.value.label), 
                                  sep = "~~~")),
               formula = y ~ x, 
               parse = TRUE, 
               size = 4) +
  labs(
    x = "Lithic Volumetric Density",
    y = "Retouch Frequency"
  ) +
  scale_x_log10() +
  scale_y_log10() +
  theme(
    axis.text = element_text(size = 7),
    axis.title = element_text(size = 10),
    plot.title = element_text(size = 12)
  )

@@

@@

sum_depth_sheet_subset <- 
 sum_depth_sheet_for_pca %>%
  drop_na(tool_diversity) #%>% 
  # filter(!label %in% c("tp1_c_3_a",
  #                      "d1_a_2_b",
  #                     "d1_c_3_b"))

@@

contribution_plots <- 
map(1:5, ~fviz_contrib(res.pca, 
                   choice = "var", 
                   axes = .x, 
                   top = 10))

#combine individual plots into one plot 
library(cowplot)
plot_grid(plotlist = contribution_plots,
                   ncol = 2)
```

@@

eigen_df <- data.frame(
   eigenvalue = res.pca$eig[,1],
  `percentage of variance` = res.pca$eig[,2],
  `cumulative percentage of variance` = res.pca$eig[,3]
)

@@

# Re-order the columns based on the corrected names
eigen_df <- eigen_df[, c("Dimension", 
                         "eigenvalue", 
                         "percentage.of.variance",
                         "cumulative.percentage.of.variance")]

@@

# Removing row names
rownames(eigen_df) <- NULL

@@

# Generate the base biplot
biplot_base <- fviz_pca_biplot(res.pca,
                               axes = c(1, 2),
                               labelsize = 2,
                               label = "var",
                               addEllipses = TRUE,
                               col.ind = excavation_area
                              )

@@

# Add geom_text_repel() to the base plot
biplot_modified <- biplot_base +
                   geom_text_repel(data = pca_label_text,
                                   aes(x = Dim.1, 
                                       y = Dim.2, 
                                       label = labels,
                                       colour = area),
                                   segment.color = "grey80",
                                   size = 2,
                                   force = 10
                                  )

@@

>>>>

```{r}
# Read the data
tl <- read_csv(here("data/raw-data/Tulai_Lithic_Assamblage.csv"))

# this file is prepared from the google sheet here
# https://docs.google.com/spreadsheets/d/1Q0QZESk412ZQLE24yPs6Rg-7Y9OMX63DHoXeIPNAIFM/edit#gid=0
# and simplified using code in the scraps.R file
excavation_data <- read_csv(here("data//raw-data/excavation_data.csv"))
```

@@

# BM: export this table as a CSV to paste into the 
# drafting document to prepare for submission

write_csv(summary_table_area_filtered,
          here("data/derived-data/tbl-001-summary_table_area_filtered.csv"))

# BM: this is table 1 in the manuscript
knitr::kable(summary_table_area_filtered,
             caption = "Summary of lithics in the main areas")
```

@@

# BM: export this table as a CSV to paste into the 
# drafting document to prepare for submission

write_csv(summary_table_tool_area,
          here("data/derived-data/tbl-002-summary_table_tool_area.csv"))

@@

# Create the plot: WABI_Level_logscale
# here I do not consider those levels with a value 0 of retouch frequency. 
filtered_data <- 
  sum_depth_sheet %>% 
  filter(lithic_density > 0 & retouch_frequency > 0)  %>% 
  mutate(excavation_area =  str_extract(label2, "[^_]+"))

WABI_Level <- ggplot(filtered_data,  
            aes(x = lithic_density,
                y = retouch_frequency)) +
  geom_point(aes(color = excavation_area,
                 shape = excavation_area), 
             size = 4) +
  geom_smooth(method = "lm", 
              se = FALSE, 
              color = "blue", 
              linewidth = 0.5) +
  # Add text labels based on 'label' column
  # BM: this is useful for investigating and
  # preparing, but not really informative 
  # for the reader, so we omit for publication
  # geom_text_repel(aes(label = label2)) +
  stat_poly_eq(aes(label = paste(stat(eq.label), 
                                  stat(rr.label), 
                                  stat(p.value.label), 
                                  sep = "~~~")),
               formula = y ~ x, 
               parse = TRUE, 
               size = 5,
               label.x = "right",
               label.y = "bottom") +
  labs(
    x = "Lithic Volumetric Density",
    y = "Retouch Frequency",
    colour = "Excavation area",
    shape = "Excavation area"
  ) +
  scale_x_log10() +
  scale_y_log10() +
  theme_minimal() +
  theme(
    legend.position = c(.8,.3),
    axis.text = element_text(size = 7),
    axis.title = element_text(size = 10),
    plot.title = element_text(size = 12)
  )

@@

# BM: save this figure to paste into the 
# drafting document to prepare for submission

ggsave(here("figures/fig-wabi-plot.jpg"),
       h = 6, 
       w = 8)

@@

sum_depth_sheet_subset <- 
 sum_depth_sheet_for_pca %>%
  drop_na(tool_diversity) %>% 
  # explore the effect of outliers on the PCA space
 filter(!label2 %in% c("tp1_b_1_120_140",
                        "tp1_c_3_0_10"))

@@

contribution_plots <- 
map(1:4, ~fviz_contrib(res.pca, 
                   choice = "var", 
                   axes = .x, 
                   top = 10))

#combine individual plots into one plot 
library(cowplot)
plot_grid(plotlist = contribution_plots,
                   ncol = 2)

# BM: save this figure to paste into the 
# drafting document to prepare for submission

ggsave(here("figures/fig-contribution-plots.jpg"),
       h = 6, 
       w = 8)

@@

eigen_df <- tibble(
   Eigenvalue = res.pca$eig[,1],
  `Percentage of Variance` = res.pca$eig[,2],
  `Cumulative Percentage of Variance` = res.pca$eig[,3]
)

@@

# Re-order the columns based on the corrected names
eigen_df <- eigen_df[, c("Dimension", 
                         "Eigenvalue", 
                         "Percentage of Variance",
                         "Cumulative Percentage of Variance")]

@@

# BM: export this table as a CSV to paste into the 
# drafting document to prepare for submission

write_csv(eigen_df,
          here("data/derived-data/tbl-003-pca-eigenvalues.csv"))

@@

# Generate the base biplot
biplot_base <- fviz_pca_biplot(res.pca,
                               axes = c(1, 2),
                               labelsize = 2,
                               label = "var",
                               addEllipses = TRUE,
                               col.ind = excavation_area,
                               title = ""
                              )

@@

# Add geom_text_repel() to the base plot
biplot_modified <- biplot_base +
                   geom_text_repel(data = pca_label_text,
                                   aes(x = Dim.1, 
                                       y = Dim.2, 
                                       label = labels,
                                       colour = area),
                                   segment.color = "grey80",
                                   size = 2,
                                   force = 10
                                  ) +
  labs(colour = "",
       shape = "",
       fill = "") +
  theme(legend.position = c(0.9, 0.2))

@@

# BM: save this figure to paste into the 
# drafting document to prepare for submission

ggsave(here("figures/fig-pca-biplot.jpg"),
       h = 6, 
       w = 8)
